### 2015-05-02 paper.tex
initial checkin

% Template for ICIP-2013 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{spconf,amsmath,graphicx}

% Example definitions.
% --------------------
\def\x{{\mathbf x}}
\def\L{{\cal L}}

% Title.
% ------
\title{Investigation on Parallelization of Deep Neural Network Training using Multiple GPUs}
%
% Single address.
% ---------------
\name{Hang Su$^1$$^2$, Haoyu Chen$^1$, Nelson Morgan$^2$}
\address{$^1$ International Computer Science Institute, Berkeley, California, US \\
$^2$ Dept. of Electrical Engineering \& Computer Science, University of California, Berkeley, CA, USA \\
{\small \tt \{suhang3240@gmail.com, williamchenthu@gmail.com\}}
}
%
% For example:
% ------------
%\address{School\\
%	Department\\
%	Address}
%
% Two addresses (uncomment and modify for two-address case).
% ----------------------------------------------------------
%\twoauthors
%  {A. Author-one, B. Author-two\sthanks{Thanks to XYZ agency for funding.}}
%	{School A-B\\
%	Department A-B\\
%	Address A-B}
%  {C. Author-three, D. Author-four\sthanks{The fourth author performed the work
%	while at ...}}
%	{School C-D\\
%	Department C-D\\
%	Address C-D}
%
\begin{document}
%\ninept
%
\maketitle
%
\begin{abstract}
In this paper we introduce Butterfly mixing to parallel training of deep neural networks (DNN). Parallelization is done
in a model averaging manner. Data is partitioned and distributed to different nodes for local model update, and
model averaging (reduce) is done every few minibatches. We compare several different reduce stratgies, including
all reduce, butterfly mixing, ring reduce and hopping ring reduce. We show that all these methods can effectively
speed up neural network training. On swithboard data, a xx times speed up is achieved using 16 gpus.

\end{abstract}
%
\begin{keywords}
Parallel training, multiple GPUs, neural network, butterfly mixing
\end{keywords}
%
\section{Introduction}
\label{sec:intro}
Deep Neural Networks (DNN) has shown its effeciveness in several machine learning tasks, espencially in speech
recognition. The large model size and massive training examples make DNN a powerful model for classification. However,
these two factors also slow down the training procedure of DNNs.

Parallelization of DNN training has been a popular topic since the revive of neural networks. Several different strategies
have been proposed to tackle this problem. Multiple thread CPU parallelization and single CUDA GPU implementation are compared
in \cite{scanzio2010parallel,vesely2010parallel}, and they show that single GPU could beat 8 cores CPU by a factor of 2.

Optimality for parallelization of DNN training was analyzed in \cite{seide2014parallelizability}, and based on the analysis, 
a gradient quantization approach was proposed to minimize communication cost \cite{seide20141}.

DistBelief proposed in \cite{dean2012large} reports a speed up of 2.2x using 8 CPU coresthan using a
single machine.

Asynchronous SGD using multiple GPUs achieved a 3.2x speed-up on 4 GPUs \cite{zhang2013asynchronous}.

A pipeline training approach was propoased in \cite{chen2012pipelined} and a 3.3x speedup was achieved using 4 GPUs, but this
method does not scale beyond number of layers in the neural network.

A speedup of 6x to 14x was achieved using 16 GPUs on training convolutional neural networks \cite{coates2013deep}. In this approach,
each GPU is responsible for a partition of the neural network. This approach is more useful for image classification where 
local structure of the neural network could be exploited.

Distributed model averaging using CPUs is proposed in \cite{zhang2014improving},
and a further improvement is done using natural gradient \cite{povey2014parallel}.

Butterfly mixing was proposed in \cite{zhao2013butterfly} to interleave communication with computation.

\section{Data parallelization and Model Averging}

\section{Different Reduce Strategies}

\subsection{Butterfly Mixing}

All printed material, including text, illustrations, and charts, must be kept
within a print area of 7 inches (178 mm) wide by 9 inches (229 mm) high. Do
not write or print anything outside the print area. The top margin must be 1
inch (25 mm), except for the title page, and the left margin must be 0.75 inch
(19 mm).  All {\it text} must be in a two-column format. Columns are to be 3.39
inches (86 mm) wide, with a 0.24 inch (6 mm) space between them. Text must be
fully justified.

\subsection{Ring \& Hopping Ring Reduce}

The paper title (on the first page) should begin 1.38 inches (35 mm) from the
top edge of the page, centered, completely capitalized, and in Times 14-point,
boldface type.  The authors' name(s) and affiliation(s) appear below the title
in capital and lower case letters.  Papers with multiple authors and
affiliations may require two or more lines for this information. Please note
that papers should not be submitted blind; include the authors' names on the
PDF.

\section{Experimental Results}
\subsection{Reduce Frequency}

\subsection{Reduce type}

\subsection{Scaling factor}

\section{Conclusion}

\section{Acknowledgements}
We would like to thank Forrest Iandola for helpful suggestion.

% References should be produced using the bibtex program from suitable
% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
% style file from IEEE produces unsorted bibliography list.
% -------------------------------------------------------------------------
\bibliographystyle{IEEEbib}
\bibliography{paper}

\end{document}

### 2015-06-15 paper.tex
paper.tex and paper.bib are not executables
add some figures

@@
 % ------
-\title{Investigation on Parallelization of Deep Neural Network Training using Multiple GPUs}
+\title{Parallel Training of Deep Neural Network using Model Averaging and Butterfly Mixing}
 %
@@
 % ---------------
-\name{Hang Su$^1$$^2$, Haoyu Chen$^1$, Nelson Morgan$^2$}
+\name{Hang Su$^{1,2}$, Haoyu Chen$^1$, Haihua Xu$^3$}
 \address{$^1$ International Computer Science Institute, Berkeley, California, US \\
 $^2$ Dept. of Electrical Engineering \& Computer Science, University of California, Berkeley, CA, USA \\
-{\small \tt \{suhang3240@gmail.com, williamchenthu@gmail.com\}}
+$^3$ Nanyang Technological University, Singapore \\
+{\small \tt \{suhang3240@gmail.com\}}
 }
@@
 \begin{abstract}
-In this paper we introduce Butterfly mixing to parallel training of deep neural networks (DNN). Parallelization is done
-in a model averaging manner. Data is partitioned and distributed to different nodes for local model update, and
-model averaging (reduce) is done every few minibatches. We compare several different reduce stratgies, including
-all reduce, butterfly mixing, ring reduce and hopping ring reduce. We show that all these methods can effectively
-speed up neural network training. On swithboard data, a xx times speed up is achieved using 16 gpus.
+In this paper we apply model averaging and butterfly mixing to parallel training of deep neural network (DNN). 
+Parallelization is done in a model averaging manner. Data is partitioned and distributed to different nodes for 
+local model updates, and model averaging (reduce) is done every few minibatches.
+
+We use multiple GPUs for data parallelization, and Message Passing Interface (MPI) for communication between nodes. 
+Natrual Gradient for Stochasitc Gradient Descent (NG-SGD) approach is used to ensure better convergence in this framework. On the
+300h Swithboard dataset, a 12 times speed up is achieved using 16 gpus with no decoding accuracy loss. We also 
+compare all-reduce strategy with butterfly mixing, and show that butterfly mixing can further save communication costs.
 
@@
 \begin{keywords}
-Parallel training, multiple GPUs, neural network, butterfly mixing
+Parallel training, model averaging, deep neural network, butterfly mixing
 \end{keywords}
@@
 
-Distributed model averaging using CPUs is proposed in \cite{zhang2014improving},
-and a further improvement is done using natural gradient \cite{povey2014parallel}.
+Distributed model averaging using CPUs is proposed in \cite{zhang2014improving}, and a further improvement 
+is done using natural gradient \cite{povey2014parallel}. Our approach is mainly based on this idea, and the main 
+difference between our work and this one is that we are performing neural networks training via MPI. Also,
+we introduce butterfly-mixing strategy for model averaging.
 
-Butterfly mixing was proposed in \cite{zhao2013butterfly} to interleave communication with computation.
+Butterfly mixing was proposed in \cite{zhao2013butterfly} to interleave communication with computation. In each
+reduce period, information (gradient or model) is exchanged between each node and its friend node. The pair of nodes
+are shuffled after every reduce so that information in each node can propagate to all other nodes in 
+$log(numNodes)$ periods.
+
+In Section 2 we introduce the model approach. In Section 3 we introduce different reduce strategies. Section 4 
+explains NG-SGD. Section 5 records experimental results on different setups and Section 6 concludes.
 
 \section{Data parallelization and Model Averging}
+Stochastic gradient descent is a popular method for DNN training, even though this is a non-convex optimization 
+problem. Since the size of training data is large, minibatch based SGD is often used for model training. 
+Roughly, the larger the minibatch size, the higher the converge rate. However, a large minibatch requires
+more compute and memory usage, so a straight forward idea for parallelization would be distributing the 
+gradient computation to different computing nodes. In each step, the gradient is reduced to one computing node,
+averaged and then used to update models in each node. This method can compute the gradient accurately, but 
+it requires heavy communication between nodes.
 
-\section{Different Reduce Strategies}
+However, if we choose to reduce the weight rather than gradient, it is not necessary to communicate that often. There
+is straight forward theory that guarantees convergence, but we do observe reasonable results in experiments.
 
-\subsection{Butterfly Mixing}
+\section{Model Averaging Strategy}
+\subsection{All-reduce}
+All-reduce strategy collects the weights from all the nodes in the network. The communication is bounded by 
+the bandwidth. Fig~\ref{fig:allreduce} is an example of all-reduce with 4 nodes. One node is responsible for
+collecting all the model parameters. It will average the weights and then broadcast it to all the nodes. 
+This approach compute the average weight based on all the information in computing nodes. Thus, we expect 
+the converge speed to be the fastest.
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.42\textwidth]{allreduce.jpg}
+  \caption{All-reduce network}
+  \label{fig:allreduce}
+\end{figure}
 
-All printed material, including text, illustrations, and charts, must be kept
-within a print area of 7 inches (178 mm) wide by 9 inches (229 mm) high. Do
-not write or print anything outside the print area. The top margin must be 1
-inch (25 mm), except for the title page, and the left margin must be 0.75 inch
-(19 mm).  All {\it text} must be in a two-column format. Columns are to be 3.39
-inches (86 mm) wide, with a 0.24 inch (6 mm) space between them. Text must be
-fully justified.
+\subsection{Butterfly mixing}
+Butterfly mixing is a reduction strategy proposed in \cite{zhao2013butterfly}. It reduces the communication load in 
+every iteration: one node would only send and receive message from one other node. The pairing of nodes are shuffled after
+each reduction so that information collected in each nodes will spread to all other nodes. Fig~\ref{fig:butterfly} 
+is an example of butterfly mixing with 4 nodes. Butterfly mixing requires much less communication than all-reduce 
+strategy. However, each node in the network only collects information from two computing nodes at a time.
+It takes $log(numNodes)$ communication to spread the message to the whole network. Thus, the converge speed 
+of butterfly mixing would be slower than all-reduce strategy. 
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.42\textwidth]{butterfly.jpg}
+  \caption{Butterfly mixing network}
+  \label{fig:butterfly}
+\end{figure}
 
-\subsection{Ring \& Hopping Ring Reduce}
+\subsection{Data exchange analysis}
 
-The paper title (on the first page) should begin 1.38 inches (35 mm) from the
-top edge of the page, centered, completely capitalized, and in Times 14-point,
-boldface type.  The authors' name(s) and affiliation(s) appear below the title
-in capital and lower case letters.  Papers with multiple authors and
-affiliations may require two or more lines for this information. Please note
-that papers should not be submitted blind; include the authors' names on the
-PDF.
+\section{Natural Gradient for Model Update}
+In stochastic gradient descent (SGD), the learning rate is often assumed to be a scalar $\alpha_t$ that may change over time,
+the update formula for model parameters $\theta_{t}$ is
+\begin{equation}
+\theta_{t+1} = \theta_{t} + \alpha_t g_t
+\end{equation}
+where $g_t$ is the gradient.
+
+However, according to Natural Gradient idea \cite{murata1999statistical,roux2008topmoumoute}, it is possible to replace 
+the scalar with a symmetric positive definite matrix $E_t$, which is the inverse of the Fisher information matrix.
+\begin{equation}
+\theta_{t+1} = \theta_{t} + \alpha_t E_t g_t
+\end{equation}
+
+Suppose $x$ is the variable we are modeling, and $f(x;\theta)$ is the probability or likelihood of $x$ given parameters $\theta$, then the
+Fisher information matrix $I(\theta)$ is defined as
+\begin{equation}
+\frac{\partial}{\partial\theta}\log f(x;\theta)
+\end{equation}
+
+For large scale speech recognition, it is impossible to estimate Fisher information matrix and perform inversion, 
+so it is necessary to approximate the inverse Fisher information matrix directly. We do not include too much 
+detail here because this is not the main focus of this report. Details about implementation of NG-SGD could 
+be found in \cite{povey2014parallel}.
 
 \section{Experimental Results}
-\subsection{Reduce Frequency}
+\subsection{Setup}
 
-\subsection{Reduce type}
+\subsection{Switchboard}
 
-\subsection{Scaling factor}
+scaling factor
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.5\textwidth]{scaling.png}
+  \caption{Scaling factor and speedup factor v.s. number of gpus}
+  \label{fig:scaling}
+\end{figure}
 
-\section{Conclusion}
+percentage of time spent on communication
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.5\textwidth]{percent.png}
+  \caption{Percentage of time on MPI communication v.s. number of gpus}
+  \label{fig:percent}
+\end{figure}
+
+\section{Conclusion and Future Work}
+Neural network training can be efficiently speed up by using model averaging techniques and NG-SGD. 
+Butterfly-mixing and ring reduction can reduce communication time a lot, but the models may not converge as well as 
+Allreduce when the number of jobs goes up to 16.
+
+CUDA aware MPI, learning rate schedule.
 
 \section{Acknowledgements}
-We would like to thank Forrest Iandola for helpful suggestion.
+We would like to thank Karel Vesely and Daniel Povey who wrote the original "nnet1" neural network training code
+and natural gradient stochastic gradient descent upon which the work here is based. We would also like to thank Nelson 
+Morgan, Forrest Iandola and Yuansi Chen for their helpful suggestions.
 

### 2015-06-16 paper.tex
draft sent to Haihua

@@
 % ------
-\title{Parallel Training of Deep Neural Network using Model Averaging and Butterfly Mixing}
+\title{Investigation on Parallel Training of Deep Neural Network using Model Averaging}
 %
@@
 We use multiple GPUs for data parallelization, and Message Passing Interface (MPI) for communication between nodes. 
-Natrual Gradient for Stochasitc Gradient Descent (NG-SGD) approach is used to ensure better convergence in this framework. On the
-300h Swithboard dataset, a 12 times speed up is achieved using 16 gpus with no decoding accuracy loss. We also 
-compare all-reduce strategy with butterfly mixing, and show that butterfly mixing can further save communication costs.
+We investigate the effectiveness of Natrual Gradient for Stochasitc Gradient Descent (NG-SGD) in the model averaging
+framework, and compare two learning rate schedule strategies for model update. We show that combination of 
+NG-SGD and exponential learning rate schedule can lead to better convergence. On the 300h Swithboard dataset, a 10 
+times speed up is achieved using 16 gpus with little decoding accuracy loss -- 0.2 in word error rate (WER). We also 
+introduce a butterfly mixing strategy that can further save communication costs in model averaging framework.
 
@@
   \centering
-  \includegraphics[width=0.42\textwidth]{allreduce.jpg}
+  \includegraphics[width=0.5\textwidth]{allreduce.png}
   \caption{All-reduce network}
@@
   \centering
-  \includegraphics[width=0.42\textwidth]{butterfly.jpg}
+  \includegraphics[width=0.5\textwidth]{butterfly.png}
   \caption{Butterfly mixing network}
@@
 
+\section{Initialization}
+Conjecture: dbn pretraining initialize the neural network such that it lies in a convex-like area
+
 \section{Experimental Results}
 \subsection{Setup}
+
+\subsection{Learing Rate Schedule}
 
@@
 \end{figure}
+
+\begin{table}
+  \centering
+  \begin{tabular}{c|c|c|c|c|c}
+    \hline
+     Nodes  & 1    & 2    & 4    & 8    & 16 \\
+    \hline
+Allreduce &    14.9 & -- & -- & 15.1 & 15.1 \\
+    \hline
+Butterfly &    14.9 & -- & -- & -- & --\\
+    \hline
+  \end{tabular}
+  \caption{Comparison of WERs using different number of GPUs }
+  \label{tab:wer}
+\end{table}
 

### 2015-06-17 paper.tex
refine the introduction

@@
 % ------
-\title{Investigation on Parallel Training of Deep Neural Network using Model Averaging}
+\title{Parallel Training of Deep Neural Network using Model Averaging and Butterfly Mixing}
 %
@@
 NG-SGD and exponential learning rate schedule can lead to better convergence. On the 300h Swithboard dataset, a 10 
-times speed up is achieved using 16 gpus with little decoding accuracy loss -- 0.2 in word error rate (WER). We also 
-introduce a butterfly mixing strategy that can further save communication costs in model averaging framework.
-
+times speed up is achieved using 16 gpus with little decoding accuracy loss -- a 0.2\% degradation in word error rate 
+(WER). We also introduce a butterfly mixing strategy that can further save communication costs in model averaging framework.
 \end{abstract}
@@
 
-Parallelization of DNN training has been a popular topic since the revive of neural networks. Several different strategies
-have been proposed to tackle this problem. Multiple thread CPU parallelization and single CUDA GPU implementation are compared
-in \cite{scanzio2010parallel,vesely2010parallel}, and they show that single GPU could beat 8 cores CPU by a factor of 2.
+Parallelization of DNN training has been a popular topic since the revival of neural networks. Several different strategies
+have been proposed to tackle this problem. Multiple thread CPU parallelization and single GPU implementation are compared
+in \cite{scanzio2010parallel,vesely2010parallel}, and it is shown that single GPU could beat multi-threaded CPU implementation
+by a factor of 2.
 
 Optimality for parallelization of DNN training was analyzed in \cite{seide2014parallelizability}, and based on the analysis, 
-a gradient quantization approach was proposed to minimize communication cost \cite{seide20141}.
+a gradient quantization approach (1-bit SGD) was proposed to minimize communication cost \cite{seide20141}. It shows that 1 bit
+can effectively reduce the amount of data exchanged in an MPI framework, and a 10 times speed-up using 40 GPUs.
 
-DistBelief proposed in \cite{dean2012large} reports a speed up of 2.2x using 8 CPU coresthan using a
-single machine.
-
-Asynchronous SGD using multiple GPUs achieved a 3.2x speed-up on 4 GPUs \cite{zhang2013asynchronous}.
+DistBelief proposed in \cite{dean2012large} reports that 8 CPU machines train 2.2 times faster than a single GPU machine on a
+moderately sized speech model. Asynchronous SGD using multiple GPUs achieved a 3.2x speed-up on 4 GPUs \cite{zhang2013asynchronous}.
 
@@
 each GPU is responsible for a partition of the neural network. This approach is more useful for image classification where 
-local structure of the neural network could be exploited.
+local structure of the neural network could be exploited. For a fully connected speech model, a model partition approach 
+may not be able to contribute as much.
 
 Distributed model averaging using CPUs is proposed in \cite{zhang2014improving}, and a further improvement 
-is done using natural gradient \cite{povey2014parallel}. Our approach is mainly based on this idea, and the main 
-difference between our work and this one is that we are performing neural networks training via MPI. Also,
-we introduce butterfly-mixing strategy for model averaging.
+is done using natural gradient \cite{povey2014parallel}. In this approach, separate models are trained on multiple nodes using 
+different partitions of data, and model parameters are averaged after each epoch. It is shown that NG-SGD can effectively improve
+convergence and ensure a better model trained using the model averaging framework.
 
-Butterfly mixing was proposed in \cite{zhao2013butterfly} to interleave communication with computation. In each
-reduce period, information (gradient or model) is exchanged between each node and its friend node. The pair of nodes
-are shuffled after every reduce so that information in each node can propagate to all other nodes in 
-$log(numNodes)$ periods.
+Our approach is mainly based on the NG-SGD with model averaging. The main differences here in this work are:
+1. We utilize multiple GPUs in neural networks training via MPI, which allows us to perform model averaging more frequently and efficiently;
+2. We introduce a butterfly-mixing strategy for model averaging that can further speed up the training.
 
-In Section 2 we introduce the model approach. In Section 3 we introduce different reduce strategies. Section 4 
+In addition, this work simplifies the model training procedure in the following aspects:
+1. It does not require a warm-up phase where only single thread can be used for model update \cite{seide20141,povey2014parallel};
+2. It does not need to tune minibatch size to utilize parallelization power;
+3. It does not use momentum or Adagrad\cite{duchi2011adaptive} for gradient estimation.
+
+On the other hand, it requires:
+1. A proper implementation of NG-SGD, as is introduced in \cite{povey2014parallel};
+2. A bigger initial learning rate and a proper model averaging frequency;
+
+In Section 2 we introduce the model averaging approach. In Section 3 we introduce butterfly-mixing strategie. Section 4 
 explains NG-SGD. Section 5 records experimental results on different setups and Section 6 concludes.
@@
 \subsection{Butterfly mixing}
+Butterfly mixing was proposed in \cite{zhao2013butterfly} to interleave communication with computation. In each
+reduce period, information (gradient / model) is exchanged between each node and its friend node. The pair of nodes
+are shuffled after every reduce so that information in each node can propagate to all other nodes in $log(numNodes)$ periods.
+
 Butterfly mixing is a reduction strategy proposed in \cite{zhao2013butterfly}. It reduces the communication load in 

### 2015-06-17 paper.tex
add some tables

@@
 
-Distributed model averaging using CPUs is proposed in \cite{zhang2014improving}, and a further improvement 
-is done using natural gradient \cite{povey2014parallel}. In this approach, separate models are trained on multiple nodes using 
+Distributed model averaging is used in \cite{zhang2014improving,miao2014distributed}, and a further improvement 
+is done using NG-SGD \cite{povey2014parallel}. In this approach, separate models are trained on multiple nodes using 
 different partitions of data, and model parameters are averaged after each epoch. It is shown that NG-SGD can effectively improve
@@
 \section{Data parallelization and Model Averging}
-Stochastic gradient descent is a popular method for DNN training, even though this is a non-convex optimization 
-problem. Since the size of training data is large, minibatch based SGD is often used for model training. 
-Roughly, the larger the minibatch size, the higher the converge rate. However, a large minibatch requires
-more compute and memory usage, so a straight forward idea for parallelization would be distributing the 
-gradient computation to different computing nodes. In each step, the gradient is reduced to one computing node,
-averaged and then used to update models in each node. This method can compute the gradient accurately, but 
-it requires heavy communication between nodes.
+SGD is a popular method for DNN training. Even though neural network training objectives are usually non-convex, 
+mini-batch SGD has been shown to be effective for optimizing the objective\cite{seide2011conversational}. 
+Roughly speaking, a bigger minibatch size gives a better estimate of the gradient, resulting in a better the converge rate. 
+Thus, a straight forward idea for parallellization would be distributing the gradient computation to different computing
+nodes. In each step, gradients of minibatches on different nodes are reduced to a single node, averaged and then used to 
+update models in each node. This method can compute the gradient accurately, but it requires heavy communication between nodes.
+Also, it is shown that increasing minibatch size does not always benefit model training\cite{}, especially in early stage 
+of model training.
 
-However, if we choose to reduce the weight rather than gradient, it is not necessary to communicate that often. There
-is straight forward theory that guarantees convergence, but we do observe reasonable results in experiments.
+On the other hand, if we choose to average the parameters rather than gradients, it is not necessary to exchange data that often. 
+Currently, there is no straight forward theory that guarantees convergence, but we would like to explore a bit why this strategy 
+shoule work, just as we observe in the experiments.
+
+First, in the extreme case where model parameters are averaged after each minibatch update, model averaging is equivalent to 
+gradient averaging. If we view the difference of model parameters as gradient in a special space.
+
+Second, it is shown that model averging for convex models is guaranteed to converge \cite{mcdonald2010distributed,mcdonald2009efficient}.
+It is suggested that unsupervised pretraining guides the learning towards basins of attraction of minima that support better generalization
+from the training data set; \cite{erhan2010does}.
 
@@
 \end{figure}
+
+MPI implementation
 
@@
 
-\section{Initialization}
-Conjecture: dbn pretraining initialize the neural network such that it lies in a convex-like area
-
 \section{Experimental Results}
@@
 
+Reduce frequency
+
 \subsection{Learing Rate Schedule}
+Why do we need a large learning rate?
 
-\subsection{Switchboard}
+\subsection{Initialization Matters}
+Compare random init and dbn init, single and paralle
+\begin{table}
+  \centering
+  \begin{tabular}{c|c|c}
+    \hline
+     Nodes          &  1    & 16 \\
+    \hline
+    random init     & --    & --  \\
+    \hline
+    rbm pretraining & --    & --\\
+    \hline
+  \end{tabular}
+  \caption{Comparing rbm pretraining with random initialization in model averaging setup }
+  \label{tab:init}
+\end{table}
 
+\subsection{Online NG-SGD Matters}
+\begin{table}
+  \centering
+  \begin{tabular}{c|c|c}
+    \hline
+     Nodes  &  1    & 16 \\
+    \hline
+    SGD     & --    & --  \\
+    \hline
+    NG-SGD  & --    & --\\
+    \hline
+  \end{tabular}
+  \caption{Comparing NG-SGD and naive SGD}
+  \label{tab:ngsgd}
+\end{table}
+
+\subsection{Speedup Comparison}
 scaling factor

### 2015-06-24 paper.tex
adding more text and experiment results

@@
 \begin{abstract}
-In this paper we apply model averaging and butterfly mixing to parallel training of deep neural network (DNN). 
+In this paper we apply model averaging to parallel training of deep neural network (DNN). 
 Parallelization is done in a model averaging manner. Data is partitioned and distributed to different nodes for 
-local model updates, and model averaging (reduce) is done every few minibatches.
+local model updates, and model averaging across nodes is done every few minibatches.
 
@@
 We investigate the effectiveness of Natrual Gradient for Stochasitc Gradient Descent (NG-SGD) in the model averaging
-framework, and compare two learning rate schedule strategies for model update. We show that combination of 
-NG-SGD and exponential learning rate schedule can lead to better convergence. On the 300h Swithboard dataset, a 10 
-times speed up is achieved using 16 gpus with little decoding accuracy loss -- a 0.2\% degradation in word error rate 
-(WER). We also introduce a butterfly mixing strategy that can further save communication costs in model averaging framework.
+framework, and compare two learning rate schedule strategies for model update. It is shown that NG-SGD leads to better
+convergence in model averaging framework. On the 300h Swithboard dataset, a 10 times speedup is achieved using 16 GPUs
+and 17 times speedup using 32 GPUs with a small decoding accuracy loss -- 0.4\% degradation in word error rate 
+(WER) respectively. We also introduce a butterfly mixing strategy that can further save communication costs in model 
+averaging framework.
 \end{abstract}
@@
 
-Our approach is mainly based on the NG-SGD with model averaging. The main differences here in this work are:
-1. We utilize multiple GPUs in neural networks training via MPI, which allows us to perform model averaging more frequently and efficiently;
-2. We introduce a butterfly-mixing strategy for model averaging that can further speed up the training.
-
-In addition, this work simplifies the model training procedure in the following aspects:
-1. It does not require a warm-up phase where only single thread can be used for model update \cite{seide20141,povey2014parallel};
-2. It does not need to tune minibatch size to utilize parallelization power;
-3. It does not use momentum or Adagrad\cite{duchi2011adaptive} for gradient estimation.
+Our approach is mainly based on the NG-SGD with model averaging. We utilize multiple GPUs in neural networks training via 
+MPI, which allows us to perform model averaging more frequently and efficiently. Currently, this work simplifies the model 
+training procedure in the following aspects:
+1. It does not add a warm-up phase where only single thread is used for model update \cite{seide20141,povey2014parallel};
+2. It does not tune minibatch size to utilize parallelization power;
 
@@
 update models in each node. This method can compute the gradient accurately, but it requires heavy communication between nodes.
-Also, it is shown that increasing minibatch size does not always benefit model training\cite{}, especially in early stage 
-of model training.
+Also, it is shown that increasing minibatch size does not always benefit model training\cite{seide2011conversational}, 
+especially in early stage of model training.
 
@@
 \subsection{All-reduce}
-All-reduce strategy collects the weights from all the nodes in the network. The communication is bounded by 
+All-reduce strategy collects the weights from all the nodes. The communication is bounded by 
 the bandwidth. Fig~\ref{fig:allreduce} is an example of all-reduce with 4 nodes. One node is responsible for
@@
 
-MPI implementation
+This operation could be easily implemented by MPI\_Allreduce
 
@@
 \subsection{Setup}
+In this paper, we report speech recognition results on the 300 hour Switchboard conversational telephone speech task 
+(Switchboard-1 Release 2). We use MSU-ISIP release of the Switchboard segmentations and transcriptions (date 11/26/02),
+together with the Mississippi State transcripts2 and the 30Kword lexicon released with those transcripts. 
+The lexicon contains pronunciations for all words and word fragments in the training data. We use the Hub5 ’00 data for
+evaluation. Specifically, we use the  the development set and Hub5 ’01 (LDC2002S13) data as a separate test set.
 
-Reduce frequency
+The Kaldi toolkit\cite{kaldi11} is used for speech recognition framework. Standard 13-dim PLP feature,
+together with 3-dim Kaldi pitch feature, is extracted and used for maximum
+likelihood GMM model training. Features are then transformed using LDA+MLLT before SAT training.
+After GMM training is done, a tanh-neuron DNN-HMM hybrid system is trained using the the 40-dimension 
+transformed fMLLR (also known as CMLLR \cite{gales1996generation}) feature as input and GMM-aligned senones 
+as targets. fMLLR is estimated in an EM fashion for both training data and test data. A trigram language model (LM) is trained 
+on 3M words of the training transcripts only.
+
+Work in this paper is built on top of the Kaldi nnet1 setup and the NG-SGD method introduced in nnet2 setup. 
+Details of DNN training follows Section 2.2 in \cite{vesely2013sequence}. In this paper, we use 6 hidden layers, where each 
+hidden layer has 2048 neurons with sigmoids. Input layer is 440 dimension (i.e. the context of 11 fMLLR frames), 
+and output layer is 8806 dimension. Mini-batch SGD is used for backpropagation and the minibatch is set to 1024 for all
+the experiments. By defult, DNNs are initialized with stacked restricted Boltzmann machines (RBMs) that are pretrained 
+in a greedy layerwise fashion \cite{hinton2006fast}. Comparison between random initialization and RBM-initialization 
+in model averaging framework is reported in Section~\ref{sec:init}.
+
+The server hardware used in this work is Stampede (TACC) (URL: https://portal.xsede.org/tacc-stampede). It is a Dell Linux 
+cluster provided as an Extreme Science Engineering Discovery Environment (XSEDE) digital service by the Texas Advanced 
+Computing Center (TACC). Stampede is configured with 6,400 Dell DCS Zeus compute nodes, the majority of which are configured
+with two 2.7 GHz E5-2680 Intel Xeon (Sandy Bridge) processors and one Intel Xeon Phi SE10P coprocessor. 128 of the nodes are 
+augmented with an NVIDIA K20 GPU and 8 GB of GDDR5 memory each, which we use for neural network training in this work.
+Stampede nodes run Linux 2.6.32 with batch services managed by the Simple Linux Utility for Resource Management (SLURM).
+
+\subsection{Averaging frequency}
+Due to the limitation of computing resource, we did not tune averaging frequency in our setup. It is set to 1 averaging
+per 20 minibatches by default. We would expect a tradeoff between lower averaging frequency (i.e. better speedup) and
+better training accuracy. But the optimal point is yet to be explored.
 
 \subsection{Learing Rate Schedule}
-Why do we need a large learning rate?
+Initial learning rate is increased in porportion to number of threads in this setup. The reason for this is straight forward:
+Assume we have $n$ minibatches of data for model training. When the model is trained using single thread, it gets updated $n$
+times. When data is distributed to $m$ machines, then each model gets updated $n/m$ times. Since the effect of model averaging 
+is mostly aggregating knowledge learnt from different data partition, the absolute change of model shall be compensated by 
+$m$ times.
+
+We compare two learning rate schedules in this section. The first one is the default setup used in Kaldi nnet1 (Newbob). 
+It starts with a initial learning rate of 0.32 and halves the rate when the improvement in frame accuracy on a cross-validation 
+set between two successive epochs falls below 0.5\%. The optimization terminates when the frame accuracy increases by less 
+than 0.1\%. Cross-validation is done on 10\% of the utterances that are held out from the training data.
+
+The second learing rate schedule is exponentially decaying. This method is used in \cite{senior2013empirical,povey2014parallel} 
+and is shown to be superior to performance scheduling and power scheduling. In this work, it starts with the same initial learning
+rate as the first method (Newbob), and decrease to the final learning rate (which is set to be 0.01 * initial learning rate). The
+number of epochs is set to $15$ in this task, which is set to be the same as Newbob scheduling.
+
+\begin{table}
+  \centering
+  \begin{tabular}{c|c|c|c|c}
+    \hline
+    & \multicolumn{2}{c|}{SWB}  & \multicolumn{2}{c}{CallHome} \\
+    \hline
+    Nodes       &  1      & 16    &  1    & 16 \\
+    \hline
+    Newbob      & 14.9    & 15.4  & 26.6  & 27.2 \\
+    \hline
+    exponential & 14.7    & 15.1  & 26.8  & 27.0 \\
+    \hline
+  \end{tabular}
+  \caption{Comparing learning rate schedule}
+  \label{tab:init}
+\end{table}
 
 \subsection{Initialization Matters}
-Compare random init and dbn init, single and paralle
+\label{sec:init}
+Compare random init and dbn init, single and parallell.
+
 \begin{table}
   \centering
-  \begin{tabular}{c|c|c}
+  \begin{tabular}{c|c|c|c|c}
     \hline
-     Nodes          &  1    & 16 \\
+    & \multicolumn{2}{c|}{SWB}  & \multicolumn{2}{c}{CallHome} \\
     \hline
-    random init     & --    & --  \\
+    Nodes          &  1     & 32     &   1    & 32\\
     \hline
-    rbm pretraining & --    & --\\
+    random init     & RR    & 16.4   &  RR    & 28.8 \\
+    \hline
+    rbm pretraining & 14.7  & 15.2   & 26.8   & 27.1 \\
     \hline
   \end{tabular}
-  \caption{Comparing rbm pretraining with random initialization in model averaging setup }
+  \caption{Comparing rbm pretraining with random initialization}
   \label{tab:init}
@@
   \centering
-  \begin{tabular}{c|c|c}
+  \begin{tabular}{c|c|c|c|c}
     \hline
-     Nodes  &  1    & 16 \\
+    & \multicolumn{2}{c|}{SWB}  & \multicolumn{2}{c}{CallHome} \\
     \hline
-    SGD     & --    & --  \\
+     Nodes  &  1    & 16        &   1   &  16 \\
     \hline
-    NG-SGD  & --    & --\\
+    SGD     & 14.9    & 16.3    &  26.9 & 28.3 \\
+    \hline
+    NG-SGD  & 14.7    & 15.1    &  26.8 & 27.0  \\
     \hline
@@
   \centering
-  \begin{tabular}{c|c|c|c|c|c}
+  \begin{tabular}{c|c|c|c|c|c|c}
     \hline
-     Nodes  & 1    & 2    & 4    & 8    & 16 \\
+     Nodes  & 1    & 2    & 4    & 8    & 16 & 32\\
     \hline
-Allreduce &    14.9 & -- & -- & 15.1 & 15.1 \\
+Allreduce &    14.7 & -- & -- & 15.1 & 15.1   & 15.2\\
     \hline
-Butterfly &    14.9 & -- & -- & -- & --\\
+Butterfly &    14.7 & -- & -- & --   & 15.6   & 16.0\\
     \hline
@@
 
+It is shown that this is
+
 \section{Conclusion and Future Work}
@@
 
-CUDA aware MPI, learning rate schedule.
+Further improvement might be achieved if parallel training runs on top of serial training initialization.
+CUDA aware MPI, learning rate schedule. Averaging frequency. Theory on convergence.
 
@@
 Morgan, Forrest Iandola and Yuansi Chen for their helpful suggestions.
+
+We acknowledge the Texas Advanced Computing Center (TACC) at The University of Texas at Austin for providing 
+HPC resources that have contributed to the research results reported within this paper (URL: http://www.tacc.utexas.edu).
 

### 2015-06-25 paper.tex
remove the butterfly mixing

@@
 \def\L{{\cal L}}
+\newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
 
@@
 \begin{abstract}
-In this paper we apply model averaging to parallel training of deep neural network (DNN). 
+In this work we apply model averaging to parallel training of deep neural network (DNN). 
 Parallelization is done in a model averaging manner. Data is partitioned and distributed to different nodes for 
@@
 
-We use multiple GPUs for data parallelization, and Message Passing Interface (MPI) for communication between nodes. 
+We use multiple GPUs for data parallelization, and Message Passing Interface (MPI) for communication between nodes,
+which allows us to perform model averaging frequently without losing much time on communication.
 We investigate the effectiveness of Natrual Gradient for Stochasitc Gradient Descent (NG-SGD) in the model averaging
 framework, and compare two learning rate schedule strategies for model update. It is shown that NG-SGD leads to better
-convergence in model averaging framework. On the 300h Swithboard dataset, a 10 times speedup is achieved using 16 GPUs
-and 17 times speedup using 32 GPUs with a small decoding accuracy loss -- 0.4\% degradation in word error rate 
-(WER) respectively. We also introduce a butterfly mixing strategy that can further save communication costs in model 
-averaging framework.
+convergence in model averaging framework. On the 300h Swithboard dataset, a 9.3 times speedup is achieved using 16 GPUs
+and 17 times speedup using 32 GPUs with limited decoding accuracy loss. 
 \end{abstract}
@@
 \begin{keywords}
-Parallel training, model averaging, deep neural network, butterfly mixing
+Parallel training, model averaging, deep neural network
 \end{keywords}
@@
 Our approach is mainly based on the NG-SGD with model averaging. We utilize multiple GPUs in neural networks training via 
-MPI, which allows us to perform model averaging more frequently and efficiently. Currently, this work simplifies the model 
-training procedure in the following aspects:
-1. It does not add a warm-up phase where only single thread is used for model update \cite{seide20141,povey2014parallel};
-2. It does not tune minibatch size to utilize parallelization power;
+MPI, which allows us to perform model averaging more frequently and efficiently. Unlike the other approach \cite{seide20141},
+we do not use a warm-up phase where only single thread is used for model update. (Admittedly, this might lead to further improvement). 
+In this work, we conduct a lot of experiments and compare different setups in model averaging framework.
 
-On the other hand, it requires:
-1. A proper implementation of NG-SGD, as is introduced in \cite{povey2014parallel};
-2. A bigger initial learning rate and a proper model averaging frequency;
-
-In Section 2 we introduce the model averaging approach. In Section 3 we introduce butterfly-mixing strategie. Section 4 
-explains NG-SGD. Section 5 records experimental results on different setups and Section 6 concludes.
+In Section 2, we introduce related works on NG-SGD. Section 3 describe the model averaging approach and some 
+intuition on the analysis. Section 4 records experimental results on different setups and Section 5 concludes.
 
@@
 Currently, there is no straight forward theory that guarantees convergence, but we would like to explore a bit why this strategy 
-shoule work, just as we observe in the experiments.
+should work, just as we observe in the experiments.
 
-First, in the extreme case where model parameters are averaged after each minibatch update, model averaging is equivalent to 
-gradient averaging. If we view the difference of model parameters as gradient in a special space.
+First, in the extreme case where model parameters are averaged after each weight update, model averaging is equivalent to 
+gradient averaging. Furthermore, if model averaging is done every $n$ minibatch based weight update, model update formula could
+be written as
+\begin{equation}
+\begin{split}
+\theta_{t+n} &= \theta_{t} +\sum_{i=0}^{n-1} \alpha g_{t+i}\\
+&= \theta_{t} +\sum_{i=0}^{n-1} \alpha \frac{\partial}{\partial\theta} F(x;\theta_{t+i})
+\end{split}
+\end{equation}
+\begin{equation}
+\overbar{\theta_{t+n}} = \overbar{\theta_{t}} +\sum_{i=0}^{n-1} \alpha \frac{\partial}{\partial\theta} \overbar{F}(x;\theta_{t+i})
+\end{equation}
+where $\theta$ is the model parameter and $\alpha$ is learning rate. If changes in model parameter $\theta$ is limited 
+within $n$ updates, this approach could be seen as an approximation to gradient averaging.
 
@@
 
-\section{Model Averaging Strategy}
-\subsection{All-reduce}
-All-reduce strategy collects the weights from all the nodes. The communication is bounded by 
-the bandwidth. Fig~\ref{fig:allreduce} is an example of all-reduce with 4 nodes. One node is responsible for
-collecting all the model parameters. It will average the weights and then broadcast it to all the nodes. 
-This approach compute the average weight based on all the information in computing nodes. Thus, we expect 
-the converge speed to be the fastest.
+Fig~\ref{fig:allreduce} is an example of all-reduce with 4 nodes. This operation could be easily implemented by MPI\_Allreduce.
 \begin{figure}[htb]
@@
 
-This operation could be easily implemented by MPI\_Allreduce
+\section{Natural Gradient for Model Update}
+This section introduces the idea proposed in \cite{povey2014parallel}.
 
-\subsection{Butterfly mixing}
-Butterfly mixing was proposed in \cite{zhao2013butterfly} to interleave communication with computation. In each
-reduce period, information (gradient / model) is exchanged between each node and its friend node. The pair of nodes
-are shuffled after every reduce so that information in each node can propagate to all other nodes in $log(numNodes)$ periods.
-
-Butterfly mixing is a reduction strategy proposed in \cite{zhao2013butterfly}. It reduces the communication load in 
-every iteration: one node would only send and receive message from one other node. The pairing of nodes are shuffled after
-each reduction so that information collected in each nodes will spread to all other nodes. Fig~\ref{fig:butterfly} 
-is an example of butterfly mixing with 4 nodes. Butterfly mixing requires much less communication than all-reduce 
-strategy. However, each node in the network only collects information from two computing nodes at a time.
-It takes $log(numNodes)$ communication to spread the message to the whole network. Thus, the converge speed 
-of butterfly mixing would be slower than all-reduce strategy. 
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.5\textwidth]{butterfly.png}
-  \caption{Butterfly mixing network}
-  \label{fig:butterfly}
-\end{figure}
-
-MPI implementation
-
-\subsection{Data exchange analysis}
-
-\section{Natural Gradient for Model Update}
 In stochastic gradient descent (SGD), the learning rate is often assumed to be a scalar $\alpha_t$ that may change over time,
@@
 For large scale speech recognition, it is impossible to estimate Fisher information matrix and perform inversion, 
-so it is necessary to approximate the inverse Fisher information matrix directly. We do not include too much 
-detail here because this is not the main focus of this report. Details about implementation of NG-SGD could 
-be found in \cite{povey2014parallel}.
+so it is necessary to approximate the inverse Fisher information matrix directly. Details about the theory and 
+implementation of NG-SGD could be found in \cite{povey2014parallel}.
 
@@
 \subsection{Setup}
-In this paper, we report speech recognition results on the 300 hour Switchboard conversational telephone speech task 
+In this work, we report speech recognition results on the 300 hour Switchboard conversational telephone speech task 
 (Switchboard-1 Release 2). We use MSU-ISIP release of the Switchboard segmentations and transcriptions (date 11/26/02),
@@
 Work in this paper is built on top of the Kaldi nnet1 setup and the NG-SGD method introduced in nnet2 setup. 
-Details of DNN training follows Section 2.2 in \cite{vesely2013sequence}. In this paper, we use 6 hidden layers, where each 
+Details of DNN training follows Section 2.2 in \cite{vesely2013sequence}. In this work, we use 6 hidden layers, where each 
 hidden layer has 2048 neurons with sigmoids. Input layer is 440 dimension (i.e. the context of 11 fMLLR frames), 
@@
 
+\subsection{Switchboard Results}
+scaling factor
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.5\textwidth]{scaling.png}
+  \caption{Scaling factor and speedup factor v.s. number of gpus}
+  \label{fig:scaling}
+\end{figure}
+
+percentage of time spent on communication
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.5\textwidth]{percent.png}
+  \caption{Percentage of time on MPI communication v.s. number of gpus}
+  \label{fig:percent}
+\end{figure}
+
+\begin{table}
+  \centering
+  \begin{tabular}{c|c|c|c|c|c|c}
+    \hline
+     Nodes  & 1    & 2    & 4    & 8    & 16 & 32\\
+    \hline
+Allreduce &    14.7 & -- & -- & 15.1 & 15.1   & 15.2\\
+    \hline
+Butterfly &    14.7 & -- & -- & --   & 15.6   & 16.0\\
+    \hline
+  \end{tabular}
+  \caption{Comparison of WERs using different number of GPUs }
+  \label{tab:wer}
+\end{table}
+
+It is shown that this is
+
+\subsection{Initialization Matters}
+\label{sec:init}
+Compare random init and dbn init.
+
+\begin{table}
+  \centering
+  \begin{tabular}{c|c|c|c|c}
+    \hline
+    & \multicolumn{2}{c|}{SWB}  & \multicolumn{2}{c}{CallHome} \\
+    \hline
+    Nodes          &  1     & 32     &   1    & 32\\
+    \hline
+    random init     & RR    & 16.4   &  RR    & 28.8 \\
+    \hline
+    rbm pretraining & 14.7  & 15.2   & 26.8   & 27.1 \\
+    \hline
+  \end{tabular}
+  \caption{Comparing rbm pretraining with random initialization}
+  \label{tab:init}
+\end{table}
+
 \subsection{Averaging frequency}
-Due to the limitation of computing resource, we did not tune averaging frequency in our setup. It is set to 1 averaging
-per 20 minibatches by default. We would expect a tradeoff between lower averaging frequency (i.e. better speedup) and
-better training accuracy. But the optimal point is yet to be explored.
+Averaging frequency here is defined as the number of minibatch-SGD performed per model averaging.
+Due to the limitation of computing resource, we only did preliminary experiments on this. Minibatch size of 1024 
+is set as default, and we compare averaging frequency of 10 and 20. It is shown that an averaging frequency of 10 gives
+slight worse speedup but a better decoding WER. The tradeoff between lower averaging frequency (i.e. better speedup) and
+better training accuracy is within expectation in that frequent model averaging means steady gradient estimation.
+
+\begin{table}
+  \centering
+  \begin{tabular}{c|c|c|c}
+    \hline
+    frequency   & Speedup   & SWB   & CallHome \\
+    \hline
+    baseline    &   --      & 14.7  & 26.8\\
+    \hline
+    10          & 9.32      & 15.1  & 27.0 \\
+    \hline
+    20          & 10.07     & 15.8  & 28.0 \\
+    \hline
+  \end{tabular}
+  \caption{Comparing different averaging frequencies}
+  \label{tab:init}
+\end{table}
+
+\subsection{Minibatch Size}
+\begin{table}
+  \centering
+  \begin{tabular}{c|c|c|c|c|c}
+    \hline
+    & \multicolumn{2}{c|}{SWB}    & \multicolumn{2}{c}{CallHome} & Speedup\\
+    \hline
+    nodes       &  1    &   16    &    1       &   16   &    \\
+    \hline
+    256         &  --      & 15.1  & 27.0     &   & \\
+    \hline
+    1024        & 9.32     & 15.1  & 27.0     &   & \\
+    \hline
+  \end{tabular}
+  \caption{Comparing different minibatch size}
+  \label{tab:init}
+\end{table}
 
@@
 
-\subsection{Initialization Matters}
-\label{sec:init}
-Compare random init and dbn init, single and parallell.
-
-\begin{table}
-  \centering
-  \begin{tabular}{c|c|c|c|c}
-    \hline
-    & \multicolumn{2}{c|}{SWB}  & \multicolumn{2}{c}{CallHome} \\
-    \hline
-    Nodes          &  1     & 32     &   1    & 32\\
-    \hline
-    random init     & RR    & 16.4   &  RR    & 28.8 \\
-    \hline
-    rbm pretraining & 14.7  & 15.2   & 26.8   & 27.1 \\
-    \hline
-  \end{tabular}
-  \caption{Comparing rbm pretraining with random initialization}
-  \label{tab:init}
-\end{table}
-
 \subsection{Online NG-SGD Matters}
@@
 
-\subsection{Speedup Comparison}
-scaling factor
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.5\textwidth]{scaling.png}
-  \caption{Scaling factor and speedup factor v.s. number of gpus}
-  \label{fig:scaling}
-\end{figure}
-
-percentage of time spent on communication
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.5\textwidth]{percent.png}
-  \caption{Percentage of time on MPI communication v.s. number of gpus}
-  \label{fig:percent}
-\end{figure}
-
-\begin{table}
-  \centering
-  \begin{tabular}{c|c|c|c|c|c|c}
-    \hline
-     Nodes  & 1    & 2    & 4    & 8    & 16 & 32\\
-    \hline
-Allreduce &    14.7 & -- & -- & 15.1 & 15.1   & 15.2\\
-    \hline
-Butterfly &    14.7 & -- & -- & --   & 15.6   & 16.0\\
-    \hline
-  \end{tabular}
-  \caption{Comparison of WERs using different number of GPUs }
-  \label{tab:wer}
-\end{table}
-
-It is shown that this is
-
 \section{Conclusion and Future Work}
@@
 We acknowledge the Texas Advanced Computing Center (TACC) at The University of Texas at Austin for providing 
-HPC resources that have contributed to the research results reported within this paper (URL: http://www.tacc.utexas.edu).
+HPC resources that have contributed to the research results reported within this work (URL: http://www.tacc.utexas.edu).
 

### 2015-06-25 paper.tex
more edits

@@
 
+\section{Relationship to Prior Works}
+To avoid confusion, we should mention that Kaldi\cite{kaldi11} contains two neural network recipes. The first implementation
+\footnote{Location in code: src/\{nnet,nnetbin\}} is described in \cite{vesely2013sequence} which supports Restricted Boltzmann Machine 
+pretraining \cite{hinton2006fast} and sequence-discriminative training \cite{povey2008boosted}. It uses single GPU 
+for SGD training. The second implementation \footnote{Location in code: src/{nnet2,nnet2bin}}
+\cite{zhang2014improving} was originally designed to support parallel training on multiple CPUs. Now it also supports 
+multiple GPUs for training using model averaging. By default, it uses layer-wise discriminative pretraining.
+
+Our work extends the first implementation so that it can utilize multiple GPUs using model averaging. We use MPI 
+in implementation, so file I/O is avoided during model averaging. This allows us to perform model averaging much
+more frequently.
+
 \section{Data parallelization and Model Averging}
@@
 nodes. In each step, gradients of minibatches on different nodes are reduced to a single node, averaged and then used to 
-update models in each node. This method can compute the gradient accurately, but it requires heavy communication between nodes.
+update models in each node. This method, i.e. gradient averaging, can compute the gradient accurately, but it requires 
+heavy communication between nodes.
 Also, it is shown that increasing minibatch size does not always benefit model training\cite{seide2011conversational}, 
@@
 \subsection{Switchboard Results}
-scaling factor
+Fig.~\ref{fig:scaling} shows the scaling factor and speedup plot for model averaging experiments. As is shown in the graph,
+a speedup of 17 could be achieved when 32 GPUs are used.
 \begin{figure}[htb]
@@
 \end{figure}
-
-percentage of time spent on communication
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.5\textwidth]{percent.png}
-  \caption{Percentage of time on MPI communication v.s. number of gpus}
-  \label{fig:percent}
-\end{figure}
-
+Table~\ref{tab:wer} shows the main decoding results for DNNs trained using different number of GPUs. In general, decoding
+results of DNNs trained model averaging degrades 0.3\~0.4 WER, depending on the number of GPUs used.
 \begin{table}
@@
     \hline
-Allreduce &    14.7 & -- & -- & 15.1 & 15.1   & 15.2\\
-    \hline
-Butterfly &    14.7 & -- & -- & --   & 15.6   & 16.0\\
+    SWB     & 14.7 & 14.9 & --  & 15.1   & 15.1   & 15.2\\
+    CallHM  & 26.8 & 26.8 & --  & 27.4   & 27.0   & 27.1 \\
     \hline
   \end{tabular}
-  \caption{Comparison of WERs using different number of GPUs }
+  \caption{Comparison of WERs using different number of GPUs}
   \label{tab:wer}
@@
 
-It is shown that this is
-
 \subsection{Initialization Matters}
 \label{sec:init}
-Compare random init and dbn init.
+Table~\ref{tab:init} compares random initialization with Restricted Boltzmann Machine (RBM) based initialization.
 
@@
     \hline
-    random init     & RR    & 16.4   &  RR    & 28.8 \\
+    random init     & --    & 16.4   &  --    & 28.8 \\
     \hline
-    rbm pretraining & 14.7  & 15.2   & 26.8   & 27.1 \\
+    RBM pretraining & 14.7  & 15.2   & 26.8   & 27.1 \\
     \hline
   \end{tabular}
-  \caption{Comparing rbm pretraining with random initialization}
+  \caption{Comparing RBM pretraining with random initialization}
   \label{tab:init}
 \end{table}
+As we can see in the table, random initialization is worse than RBM pretraining based DNN by xx in single GPU case.
+In model averaging setup, random initialization becomes even worse -- a xx point more degradation on WER.
 
@@
 Due to the limitation of computing resource, we only did preliminary experiments on this. Minibatch size of 1024 
-is set as default, and we compare averaging frequency of 10 and 20. It is shown that an averaging frequency of 10 gives
-slight worse speedup but a better decoding WER. The tradeoff between lower averaging frequency (i.e. better speedup) and
-better training accuracy is within expectation in that frequent model averaging means steady gradient estimation.
+is set as default, and we compare averaging frequency of 10 and 20. It is shown in Table~\ref{tab:freq} that an averaging 
+frequency of 10 gives slight worse speedup but a better decoding WER. The tradeoff between lower averaging frequency 
+(i.e. better speedup) and better training accuracy is within expectation in that frequent model averaging means 
+steady gradient estimation.
 
@@
   \caption{Comparing different averaging frequencies}
-  \label{tab:init}
+  \label{tab:freq}
 \end{table}
@@
 \subsection{Minibatch Size}
+Table~\ref{tab:mbsize} compares two different minibatch size in model averaging setup. It is shown that.
 \begin{table}
@@
     \hline
-    & \multicolumn{2}{c|}{SWB}    & \multicolumn{2}{c}{CallHome} & Speedup\\
+    & \multicolumn{2}{c|}{SWB}    & \multicolumn{2}{c|}{CallHome} & Speedup\\
     \hline
@@
     \hline
-    256         &  --      & 15.1  & 27.0     &   & \\
+    256         &  --      &      &  --      &        & \\
     \hline
@@
   \caption{Comparing different minibatch size}
-  \label{tab:init}
+  \label{tab:mbsize}
 \end{table}
@@
 \subsection{Learing Rate Schedule}
-Initial learning rate is increased in porportion to number of threads in this setup. The reason for this is straight forward:
+Initial learning rate is increased in porportion to number of threads in model averaging setup. The reason for this is straight 
+forward:
 Assume we have $n$ minibatches of data for model training. When the model is trained using single thread, it gets updated $n$
@@
 
+As is shown in Table~\ref{tab:lr}, exponential learning rate gives slight better decoding results than Newbob.
 \begin{table}
@@
   \caption{Comparing learning rate schedule}
-  \label{tab:init}
+  \label{tab:lr}
 \end{table}
@@
 \subsection{Online NG-SGD Matters}
+Table~\ref{tab:ngsgd} compares plain SGD with NG-SGD in model averaging mode, and it shows NG-SGD with 
 \begin{table}
@@
 \section{Conclusion and Future Work}
-Neural network training can be efficiently speed up by using model averaging techniques and NG-SGD. 
-Butterfly-mixing and ring reduction can reduce communication time a lot, but the models may not converge as well as 
-Allreduce when the number of jobs goes up to 16.
+In this paper, we show that neural network training can be efficiently speeded up using model averaging. 
+on a 300h Switchboard dataset, a 9.3x / 17x speedup could be achieved using 16 / 32 GPUs respectively, 
+with limited decoding accuracy loss. We also show that model averaging benefits a lot from NG-SGD and RBM based pretraining.
+Preliminary experiments on minibatch size, averaging frequency and learning rate schedules are also presented.
 
-Further improvement might be achieved if parallel training runs on top of serial training initialization.
-CUDA aware MPI, learning rate schedule. Averaging frequency. Theory on convergence.
+Further accuracy improvement might be achieved if parallel training runs on top of serial training initialization.
+It would be interesting to see if sequence-discriminative training combines well with model averaging.
+Speedup factor could be further improved if CUDA aware MPI is used. Theory on convergence using model averaging is 
+to be explored, which might be useful for guiding future development.
 
@@
 % style file from IEEE produces unsorted bibliography list.
+
 % -------------------------------------------------------------------------

### 2015-07-05 paper.tex
prepare to submit to arXiv

@@
 \usepackage{spconf,amsmath,graphicx}
+\usepackage{slashbox,pict2e}
 
@@
 % ------
-\title{Investigation on Parallel Training of Deep Neural Network using Model Averaging}
+\title{Experiments on Parallel Training of Deep Neural Network using Model Averaging}
 %
@@
 which allows us to perform model averaging frequently without losing much time on communication.
-We investigate the effectiveness of Natrual Gradient for Stochasitc Gradient Descent (NG-SGD) in the model averaging
-framework, and compare two learning rate schedule strategies for model update. It is shown that NG-SGD leads to better
-convergence in model averaging framework. On the 300h Swithboard dataset, a 9.3 times speedup is achieved using 16 GPUs
-and 17 times speedup using 32 GPUs with limited decoding accuracy loss. 
+We investigate the effectiveness of Natrual Gradient Stochasitc Gradient Descent (NG-SGD) and Restricted Boltzmann 
+Machine (RBM) pretraining for parallel training in model-averaging framework, and explore the best setups in 
+term of different learning rate schedules, averaging frequencies and minibatch sizes. It is shown that NG-SGD 
+and RBM pretraining benefits parameter-averaging based model training. On the 300h Swithboard dataset, a 9.3 times 
+speedup is achieved using 16 GPUs and 17 times speedup using 32 GPUs with limited decoding accuracy loss. 
+\footnote{This work is not submitted to 
+peer-review conferences because the authors think it needs more investigation. The authors are in lack of 
+resources to perform further exploration. However, we welcome any comments and suggestions.}
 \end{abstract}
@@
 \begin{keywords}
-Parallel training, model averaging, deep neural network
+Parallel training, model averaging, deep neural network, natural gradient
 \end{keywords}
@@
 recognition. The large model size and massive training examples make DNN a powerful model for classification. However,
-these two factors also slow down the training procedure of DNNs.
+these two factors also slow down the training procedure.
 
@@
 a gradient quantization approach (1-bit SGD) was proposed to minimize communication cost \cite{seide20141}. It shows that 1 bit
-can effectively reduce the amount of data exchanged in an MPI framework, and a 10 times speed-up using 40 GPUs.
+quantization can effectively reduce data exchange in an MPI framework, and a 10 times speed-up is achieved using 40 GPUs.
 
@@
   \centering
-  \begin{tabular}{c|c|c|c|c|c|c}
+  \begin{tabular}{|c|c|c|c|c|c|c|}
     \hline
-     Nodes  & 1    & 2    & 4   & 8      & 16     & 32\\
+    \backslashbox{Data}{Nodes}  & 1    & 2    & 4   & 8      & 16     & 32\\
     \hline
-    SWB     & 14.7 & 14.9 & --  & 15.1   & 15.1   & 15.2\\
-    CallHM  & 26.8 & 26.8 & --  & 27.4   & 27.0   & 27.1 \\
+    SWB     & 14.7 & -- & --  & 15.1   & 15.1   & 15.2\\
+    CallHM  & 26.8 & -- & --  & 27.4   & 27.0   & 27.1 \\
+    \hline
+    SWB     & 16.1  & -- & -- & 16.4  & 16.2  & 16.4 \\
+    SWB2P3  & 21.0  & -- & -- & 21.8  & 21.7  & 21.7 \\
+    SWB-Cell & 27.4 & -- & -- & 27.3  & 27.4  & 27.8 \\
     \hline
@@
     \hline
-    random init     & --    & 16.4   &  --    & 28.8 \\
+    random init     & 15.6  & 16.4   & 27.4   & 28.8 \\
     \hline
@@
 \end{table}
-As we can see in the table, random initialization is worse than RBM pretraining based DNN by xx in single GPU case.
-In model averaging setup, random initialization becomes even worse -- a xx point more degradation on WER.
+As we can see in the table, random initialization is worse than DNN with RBM pretraining by 0.9/0.6 in single GPU case.
+While in model averaging setup, random initialization becomes even worse -- 0.3/0.9 point more degradation on WER.
 
@@
 \subsection{Minibatch Size}
-Table~\ref{tab:mbsize} compares two different minibatch size in model averaging setup. It is shown that.
+Table~\ref{tab:mbsize} compares two different minibatch size in model averaging setup.
 \begin{table}
@@
     \hline
-    256         &  --      &      &  --      &        & \\
+    256         & 15.3     & 15.6 & 26.8     &  27.3   & -- \\
     \hline
-    1024        & 9.32     & 15.1 & 27.0     &        & \\
+    1024        & 14.7     & 15.1 & 26.8     &  27.0   & 9.32\\
     \hline
@@
 
-As is shown in Table~\ref{tab:lr}, exponential learning rate gives slight better decoding results than Newbob.
+As is shown in Table~\ref{tab:lr}, these two learning rate scheduling methods give similar decoding results. However, 
+exponential learning rate might need more tuning since it requires a initial learning rate, a final learning rate and
+predefined number of epochs to train.
 \begin{table}
@@
 \subsection{Online NG-SGD Matters}
-Table~\ref{tab:ngsgd} compares plain SGD with NG-SGD in model averaging mode, and it shows NG-SGD with 
+Table~\ref{tab:ngsgd} compares plain SGD with NG-SGD in model averaging mode, and it shows NG-SGD is crucial to 
+model training with parameter-averaging.
 \begin{table}
@@
 \section{Conclusion and Future Work}
-In this paper, we show that neural network training can be efficiently speeded up using model averaging. 
+In this work, we show that neural network training can be efficiently speeded up using model averaging. 
 on a 300h Switchboard dataset, a 9.3x / 17x speedup could be achieved using 16 / 32 GPUs respectively, 
